---
title: "Intelligent Monitoring and Maintenance Prediction System for Industrial Equipment"
author: "Your Team Name"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# Load required libraries
library(tidyverse)
library(knitr)
library(kableExtra)
```

# Executive Summary

Brilliant Automation, a leader in industrial automation solutions, has engaged our team to develop an intelligent monitoring and predictive maintenance system for industrial equipment. This project will leverage advanced data analytics and machine learning to transform sensor data into actionable insights, supporting early fault detection and optimized maintenance for key machinery.

# Introduction

Brilliant Automation specializes in advanced monitoring and control systems for manufacturing and processing plants. To maximize equipment reliability and operational efficiency, the company installs high-frequency vibration and temperature sensors at critical points on key machinery, such as motors, gearboxes and bearing housings, across their clients' facilities. These sensors continuously collect data to monitor the health of essential assets including Tube Mills, Belt Conveyors, and High-Temperature Fans.

This project aims to enhance their predictive maintenance capabilities. By leveraging advanced signal processing and machine learning, our system will enable early detection of equipment issues, reduce unplanned downtime, and optimize maintenance schedules. The solution will integrate sensor exploratory data analysis, machine learning predictions, and an intuitive visualization platform to help maintenance teams make informed decisions and deliver greater value to Brilliant Automation's clients.

The focus of our analysis will be on three key pieces of industrial machinery: a Tube Mill, Belt Conveyor #8, and High-Temperature Fan #1. Our solution will integrate sensor data analysis, machine learning predictions, and an intuitive visualization platform to help maintenance teams make informed decisions about equipment upkeep and repair scheduling.

## Context and Need

Manufacturing facilities face constant challenges in maintaining equipment reliability while minimizing maintenance costs. Current maintenance practices, which often rely on fixed schedules or reactive approaches, can lead to either unnecessary maintenance or unexpected breakdowns. Modern sensor technology and data analytics offer an opportunity to revolutionize this approach through data-driven decision making.

## Core Challenges

Our project addresses several key challenges in industrial maintenance:

1. Converting complex sensor readings into meaningful maintenance indicators
2. Establishing an objective system for equipment health evaluation
3. Building transparent and reliable prediction models
4. Creating an accessible interface for maintenance personnel

## Key Goals

We aim to achieve the following:

1. Data Analysis and Understanding:
   - Map relationships across different measurement types
   - Identify patterns in equipment behavior
   - Analyze vibration signatures and their implications

2. Health Assessment Framework:
   - Create a comprehensive scoring system
   - Implement real-time health monitoring
   - Ensure compatibility with maintenance standards

3. Predictive Modeling:
   - Develop transparent prediction systems
   - Enable early fault detection
   - Provide clear reasoning for predictions

4. User Interface Development:
   - Enable live monitoring
   - Present clear status indicators
   - Facilitate historical analysis

## Project Outputs

We will deliver:

1. An interactive dashboard powered by Python and Dash/Plotly
2. A robust data analysis pipeline
3. Validated predictive models
4. Complete system documentation
5. Technical presentation and detailed report

# Technical Approach

## Data Overview

Our analysis centers on high-frequency measurements from industrial equipment:

```{r data-summary-table}
data_summary <- data.frame(
  Equipment = c("Tube Mill", "Belt Conveyor #8", "High-Temperature Fan #1"),
  Sensor_Points = c("5 locations", "3 locations", "4 locations"),
  Data_Collection = c("5-second intervals", "5-second intervals", "5-second intervals"),
  Rating_Updates = c("20-minute intervals", "20-minute intervals", "20-minute intervals")
)

kable(data_summary, caption = "Measurement System Overview") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Measurement Parameters:

Primary Sensor Measurements:

- Vibration Velocity Z – RMS velocity to assess overall vibration levels.
- Low-Frequency Acceleration Z – Indicates imbalance and misalignment.
- High-Frequency Acceleration – Sensitive to early-stage bearing faults.
- Temperature – Reflects heat buildup from friction or load.

Calculated Indicators:

- crest_factor: Measures the severity of impacts in the signal, reflecting mechanical shocks.
- kurtosis_opt: Evaluates the peakiness of the signal to detect slight impacts or bearing defects.
- rotor_balance_status: Indicates whether the rotor is unbalanced due to eccentricity or uneven mass distribution.
- alignment_status: Reflects whether there is misalignment or skew in the coupler or bearing system.
- fit_condition: Describes the precision of fit between mechanical parts like shafts and bores.
- bearing_lubrication: Judges whether lubrication is sufficient or deteriorated.
- rubbing_condition: Detects abnormal rubbing, scraping, or contact between rotating elements.
- electromagnetic_status: Evaluates motor magnetic flux variations or imbalance.
- peak_value_opt: Highlights transient or shock events based on signal peak amplitude.
- rms_0_10hz: Captures imbalance or looseness.
- rms_10_100hz: Sensitive to mechanical wear or misalignment.
- rms_1_10khz: Detects early-stage bearing damage.
- rms_10_25khz: Monitors for micro-collisions and fine cracks.
- peak_10_1000hz: Detects impact events across a wide frequency range.
- velocity_rms: Measures general vibration intensity in mm/s RMS.

## Implementation Strategy

### 1. Signal Analysis

Our approach includes:

- Spectral decomposition through FFT analysis
- Signal envelope examination
- Multi-sensor pattern analysis
- Temporal behavior study

### 2. Data Processing

We will develop:

- Time-domain feature extraction
- Frequency spectrum analysis
- Dynamic window computations
- Multi-sensor data fusion

### 3. Model Development

| Model                        | Nonlinearity Handling | Risk of Overfitting | Interpretability | Feature Engineering Needed |
|:------------------------------|:---------------------:|:------------------:|:----------------:|:--------------------------:|
| Ridge (Linear)                | Bad                   | Low                | Excellent        | Beneficial                     |
| Polynomial Regression (deg=2) | Good                    | Medium             | Excellent           | Yes       |
| Random Forest                 | Natural                  | Low-Medium         | Good             | None                     |
| Gradient Boosting             | Good                | Medium             | Good               | None                     |
| Neural Network                | Excellent              | High if small data | Bad               | None                     |


### 4. Performance Assessment

We will evaluate based on:

- Prediction reliability
- Model transparency
- Error detection accuracy
- Dashboard usability

## Implementation Challenges

- **Balancing Accuracy and Interpretability**:  
  We aim to increase model accuracy while keeping it interpretable. High-complexity models (e.g., deep learning) could achieve better performance but would sacrifice explainability, which is critical for stakeholders. Therefore, careful model selection (e.g., regularized regression, shallow trees) and feature importance analysis are needed.

- **Feature and Target Mismatch**:  
  Features are collected every 5 seconds, but targets are available only every 20 minutes. This mismatch requires designing appropriate aggregation or summarization strategies (e.g., sliding windows, statistical summaries) to bridge the time scales effectively without losing important signal.

- **Feature-to-Target Ratio**:  
  We have fewer features than targets over time. This can limit the predictive power of complex models and increase the risk of overfitting. We need to focus on feature engineering, possibly creating synthetic features (e.g., rolling averages, derivatives) to enrich the input space while maintaining model simplicity.


# Project Schedule

```{r timeline}
timeline <- data.frame(
  Stage = c("Initial Analysis", "Rating System Creation", 
            "Model Construction", "Interface Development", 
            "System Validation", "Final Documentation"),
  Time_Allocation = c("2 days", "4 days", "3 weeks", "1.5 weeks", "4 days", "1 week"),
  Key_Outputs = c("Analysis Report", "Rating Framework", 
                  "Prediction Models", "Working Dashboard", 
                  "Test Results", "Complete Documentation"),
  Dates = c(
    "Apr 30 – May 2",    # initial analysis before proposal presentation
    "May 3 – May 6",     # build rating framework before proposal draft
    "May 7 – May 28",    # model building happens early (longest phase)
    "May 29 – June 9",   # build dashboard until runnable draft
    "June 10 – June 13", # validate before final presentation
    "June 14 – June 25"  # final docs up to final submission
  )
)

kable(timeline, caption = "Development Timeline") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Information Sources

1. Equipment monitoring standards and research
2. Technical documentation for analytical tools
3. Industrial maintenance best practices
